---
title: "Project03"
author: "Gabriel Santos"
date: '2022-10-21'
output: html_document
---

## Introduction

# The question, "Which are the most valued data science skills?"


## Data Source

Data from [Kaggle.com](https://www.kaggle.com/discdiver/the-most-in-demand-skills-for-data-scientists/) 


## Libraries load
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(tm)
library(wordcloud)
library(memoise)
library(SnowballC)
library(RColorBrewer)
library(RCurl)
library(XML)
library(treemap)
```

## Data Load

### Data read from the CSV file which was uploaded to Github.
```{r}
url <- "https://raw.githubusercontent.com/GabrielSantos33/DATA607_Project_03/main/DataSkills.csv"
data_skills <-read.csv(url, stringsAsFactors = FALSE)
kable(data_skills) %>% kable_styling(bootstrap_options = "striped")
```

# Tidy and Wrangle

## Data Structure
### Structure of the dataset.
```{r}
str(data_skills)
```

## Data Types
### Removed commas in the numbers and changed the data types from character to numeric for the following columns: LinkedIn, Indeed, SimplyHired and Monster.
```{r}
data_skills$LinkedIn <- str_replace_all(data_skills$LinkedIn, ",", "") %>% as.numeric()
data_skills$Indeed <- str_replace_all(data_skills$Indeed, ",", "") %>% as.numeric()
data_skills$SimplyHired <- str_replace_all(data_skills$SimplyHired, ",", "") %>% as.numeric()
data_skills$Monster <- str_replace_all(data_skills$Monster, ",", "") %>% as.numeric()
str(data_skills)
```

## Data Subset
### Removed the rows by sub-setting and eliminating those in which LinkedIn was NA; removed "Total".
```{r}
data_skills_subset <- subset(data_skills, !is.na(LinkedIn)) %>% subset(!Keyword == "Total")
kable(data_skills_subset) %>% kable_styling(bootstrap_options = "striped")
```

## Create new columns
### Crated a new column `Total_Mention`.`Total_Mention` was calculated by adding all numbers from the 4 jobboards for each skill.
```{r}
data_skills_2 <- data_skills_subset %>% mutate(Total_Mention = LinkedIn + Indeed + SimplyHired + Monster) 
kable(data_skills_2) %>% kable_styling(bootstrap_options = "striped")
```

## Add rows
### Added rows "AI" and "Artificial Intelligence" then subtracted the overlapping skills. We assigned the values to "AI + Artificial Intelligence".
```{r}
data_skills_2[18,2:6] <- data_skills_2[16,2:6] + data_skills_2[17,2:6] - data_skills_2[18,2:6]
```

### Add new rows "NLP" and "Natural Language Processing" then subtracted the overlapping skills. We assigned the values to "NLP + Natural Language Processing".
```{r}
data_skills_2[21,2:6] <- data_skills_2[19,2:6] + data_skills_2[20,2:6] - data_skills_2[21,2:6]
```

## Remove rows
### Removed the unnecessary rows "AI", "Artificial Intelligence", "NLP" and "Natural Language Processing".
```{r}
data_skills_tidy <- data_skills_2[- c(16, 17, 19, 20), ]
```

### Using mutuate function on the data frame to generate another new column `Percentage`.Percentage was calculated by dividing the total number of each skill by the overall total of all skills.
```{r}
data_skills_tidy <- data_skills_tidy %>% mutate(Percentage = Total_Mention/sum(Total_Mention))
```

# Analysis and Visualization

## Bar Plot
### Created a bar plot to show the total frequency of each data science skill mentioned in the jobboards, ranked from highest to lowest.
```{r echo=FALSE, warning=FALSE}
data_skills_tidy %>% 
  ggplot(aes(x = reorder(Keyword, -Total_Mention), y = Total_Mention, fill = Keyword)) + 
  #x-axis are skills ordered by total frequency descending, y-axis is frequency, color filled by different skills
  geom_bar(stat = "identity") + #using a bar plot
  guides(fill = FALSE) + #no legend, since the information is redundant
  xlab("Data Science Skills") + ylab("Total Frequency") + #customize x and y axis title
  ggtitle("Data Science Skills Ranked by Popularity") + #customize title
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) #rotate the x axis by 60 degrees so each skill is shown properly
```

## Treemap
### Create a treemap to show the percentage of each data science skill in respect to total.
```{r echo=FALSE, warning=FALSE}
data_skills_tidy %>% treemap(index = "Keyword", #choosing "Keyword"" as categorical variable
                             vSize = "Percentage", #choosing "Percentage" as quantitative variable (how big is each category)
                             palette = brewer.pal(17, "Set3"), #select color palette from RColorBrewer
                             title = "Proportion of Data Science Skills", #customize title
                             fontsize.title = 14) #change font size of the title
```


## Word Cloud
### Determine the most popular words in the data set. 
```{r echo=FALSE,warning=FALSE}
Corpus<- Corpus(VectorSource(data_skills_2$Keyword))
Corpus<- tm_map(Corpus, content_transformer(tolower))
Corpus<- tm_map(Corpus, removeWords, stopwords())
Corpus<- tm_map(Corpus, stripWhitespace)
wordcloud(Corpus, min.freq = 10, colors = brewer.pal(8,"Set2"),random.order = FALSE, rot.per = .1)
```

## Conclusions

### The top five keyword/skills mentioned by post throughout the different platforms were *Analysis*, *Machine Learning*, *Statistics*, *Computer Science* and *Communication*. These keywords/skills ranked differently across platforms.

In LinkedIn, the rank was as follows: *Machine Learning*, *Analysis*, *Statistics*, *Computer Science* and *Communication*.
In Indeed and Simply Hired, the rank was as follows: *Analysis*, *Machine Learning*, *Statistics*, *Computer Science* and *Communication*.
In Monster, the rank was as follows: *Analysis*, *Statistics*, *Machine Learning*, *Communication* and *Computer Science*.